---
title: "Increment_Autism_MSEC"
author: "Rebecca Netson"
date: "7/9/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(writexl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(tibble)
library(data.table)
library(tidyquant)
library(roll)
library(zoo)
```

Loading in longitudinal autism data
```{r setup, include=FALSE}
mita.db.compl.new.data = read_excel("~/Autism/longitudinal_autism_data.xlsx", col_names = TRUE)
dim(mita.db.compl.new.data)
colnames(mita.db.compl.new.data)

mita.atec.combined.compl.new.data = mita.db.compl.new.data
colnames(mita.atec.combined.compl.new.data)
print(paste("Original ", length(unique(mita.atec.combined.compl.new.data$User_ID)), " unique subjects left", sep = ''))

mita.atec.combined.compl.new.data = mita.atec.combined.compl.new.data[mita.atec.combined.compl.new.data$Treatment == 1 | mita.atec.combined.compl.new.data$Treatment == 0,]
print(paste("After filtered out Treatment that is neither 0 nor 1: ", length(unique(mita.atec.combined.compl.new.data$User_ID)), " unique subjects left", sep = ''))

#Filtering out for screentime either 1 or 0
mita.atec.combined.compl.new.data = mita.atec.combined.compl.new.data[mita.atec.combined.compl.new.data$`TVtime quartiles zero if 50 min or less; 1 if 120min or more` == 1 | mita.atec.combined.compl.new.data$`TVtime quartiles zero if 50 min or less; 1 if 120min or more` == 0,]
print(paste("After filtered out Screentime that is neither 0 nor 1: ", length(unique(mita.atec.combined.compl.new.data$User_ID)), " unique subjects left", sep = ''))


i <- c(23:28, 11, 12, 19)
mita.atec.combined.compl.new.data[, i] <- apply(mita.atec.combined.compl.new.data[, i], 2,
                                                function(x) as.numeric(x))
sapply(mita.atec.combined.compl.new.data, class)
mita.atec.combined.compl.new.data = mita.atec.combined.compl.new.data[complete.cases(mita.atec.combined.compl.new.data),]
print(paste("After removes incomplete records IN ANY COLUMN: ", length(unique(mita.atec.combined.compl.new.data$User_ID)), " unique subjects left", sep = ''))
```

Propensity matching for Visit 1
```{r setup, include=FALSE}
mita.atec.combined.compl.new.data.visit1 = mita.atec.combined.compl.new.data[mita.atec.combined.compl.new.data$Visit == 1,]
#dim(mita.atec.combined.compl.new.data.visit1)


atec.mita.new.matched.samples = matchit(as.logical(mita.atec.combined.compl.new.data.visit1$Treatment) ~
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$BL_Scale1) +
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$BL_Scale2) +
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$BL_Scale3) +
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$BL_Scale4) +
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$BL_MSEC) +
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$Gender) +
                                            as.numeric(mita.atec.combined.compl.new.data.visit1$AgeAtBaseline),
                                          data=mita.atec.combined.compl.new.data.visit1, method = "nearest", discard = "both")

mita.atec.new.matched.combined.subset = match.data(atec.mita.new.matched.samples)
print(paste("Size of a propensity score matched set is: ", dim(mita.atec.new.matched.combined.subset), sep=''))
matched.subjects = unique(mita.atec.new.matched.combined.subset$User_ID)
mita.atec.combined.compl.new.data.reduced = mita.atec.combined.compl.new.data[mita.atec.combined.compl.new.data$User_ID %in% matched.subjects,]
old.matched.subjects = unique(mita.atec.combined.compl.new.data.reduced$User_ID)
sum(mita.atec.new.matched.combined.subset$Treatment == 1)#number of treatment 
sum(mita.atec.new.matched.combined.subset$Treatment == 0)#number of controls
  
  # Making matched reduced set
mita.atec.new.matched.combined.subset = mita.atec.combined.compl.new.data.reduced  #with 2928 controls and 2928 cases
write_xlsx(mita.atec.new.matched.combined.subset, "mita.atec.new.matched.combined.subset.xlsx")
```

Matched data is new data set 
```{r setup, include=FALSE}
df <- read_excel("~/Autism/mita.atec.new.matched.combined.subset.xlsx") #5856 unique users
df$ScreenTime = as.factor(df$`TVtime quartiles zero if 50 min or less; 1 if 120min or more`)
#df$Age_Rounded = as.integer(df$Age_Rounded)
#df$Subjects = df$User_ID
#df$Endpoint_MSEC = df$MSEC
#df$Baseline_MSEC = df$BL_MSEC
```
Creating sequential series for modeling 
```{r}
df$Visit = as.Date(df$Visit) #(treating Visits as dates)
summary(df$Visit)
```
Splitting the data frame into users by creating lists within the data frame 
```{r}
df_subset = subset(df, select = c(User_ID, Visit, MSEC, BL_MSEC, Gender, Language, Severity, ScreenTime)) #Subsetting only the columns needed for model
df_subset$difference = df_subset$MSEC - df_subset$BL_MSEC #Difference from Baseline to Endpoint of each scoring 

#Creating a nested dataframe
by_user <- df_subset %>% 
  group_by(User_ID) %>%
  nest()
#Interpolating
df_subset <- df_subset %>%
  complete(nesting(User_ID,ScreenTime), Visit)
#Computing moving average for only consecutive visits per user
df_subset <- df_subset %>%
  group_by(User_ID) %>%
  mutate(moving_average_actual = rollapplyr(difference, 2, mean, fill=NA))
df_subset <- df_subset[complete.cases(df_subset),]
df_subset %>% group_by(Visit)
head(df_subset)
df_subset$User_ID <- as.factor(df_subset$User_ID)
#Graphing incremental change 
#p <- ggplot(data = df_subset, aes(x = Visit, y = moving_average_actual))+geom_point()
#p

```



```{r}
#Finding the average of the incremental average per visit across all users
df_visits <- aggregate(moving_average_actual ~ Visit, df_subset,mean)
df_visits

Visit_time_series <- df_visits %>%
  ggplot(aes(x = Visit, y = moving_average_actual))+geom_line() + geom_ma(ma_fun = SMA, n=11)
Visit_time_series

all_user_averages = aggregate(df_subset[, "moving_average_actual"], list(df_subset$Visit), mean)
```


Without imputation - group by low or high screentime
```{r}
df_subset_low = df_subset[df_subset$ScreenTime == 0, ]
df_subset_high = df_subset[df_subset$ScreenTime == 1, ]

#for low screen time users
mean_low = df_subset_low %>%
  group_by(Visit)%>%
  summarise(avg = mean(moving_average_actual, na.rm = TRUE))
ggplot(mean_low, aes(x = Visit, y = avg))+geom_point() 
#lm.low <- lm(moving_average_actual ~ Visit, data = df_subset_low)
#plot(lm.low)
p <- ggplot(df_subset_low, aes(x = Visit, y = moving_average_actual)) + geom_line()
p


#for high screen time users 
mean_high = df_subset_high %>%
  group_by(Visit) %>%
  summarise(avg = mean(moving_average_actual, na.rm = TRUE))

q <- ggplot(df_subset_high, aes(x = Visit, y = moving_average_actual)) + geom_point()+geom_line()


#mean_average <- df_subset %>%
  group_by(Visit)%>%
  summarize(mean = mean(moving_average_actual, na.rm = TRUE ))

#all users with color for the factor of screentime 
df_subset$ScreenTime = as.factor(df_subset$ScreenTime)
ggplot(df_subset, aes(x = Visit, y = moving_average_actual, color = ScreenTime)) + geom_point()
```


```{r}
mean_high$screentime = rep(1, 12)
mean_low$screentime = rep(0,12)
not_imputed <- data.frame(mean_high$Visit, mean_high$avg, mean_low$avg, mean_low$screentime, mean_low$screentime)
not_imputed
```


Imputation for missing visit values, MSEC values and BL_MSEC values
For low screentime users
```{r}
library(mice)
df_SLMiss <- df_subset_low %>%
  complete(nesting(User_ID,ScreenTime), Visit)
df_SLMiss <- subset(df_SLMiss, select = c(Visit, User_ID, BL_MSEC, MSEC))
df_SLImp <- mice(df_SLMiss, m = 3, method = "rf") #random forest impuation

complete_df_SLimp <- complete(df_SLImp, 1) #complete imputed dataest for low screentime users 
complete_df_SLimp$difference = complete_df_SLimp$MSEC - complete_df_SLimp$BL_MSEC #difference in MSEC Scores
complete_df_SLimp <- complete_df_SLimp %>% 
  group_by(User_ID) %>%
  mutate(moving_average_actual = rollapplyr(difference, 2, mean, fill=NA)) #Moving average

 
complete_df_SLimp = aggregate(complete_df_SLimp[, "moving_average_actual"], list(complete_df_SLimp$Visit), mean, na.rm = TRUE)
plot(complete_df_SLimp$Group.1, complete_df_SLimp$moving_average_actual, na.rm = TRUE)
ggplot(complete_df_SLimp, aes(x = Group.1, y = moving_average_actual)) + geom_point(na.rm = FALSE) + geom_line() + xlab("Visit") + ylab("Moving Average") + scale_x_continuous(limits = (c(1, 13)))
```

Imputation for missing visit values, MSEC values and BL_MSEC values 
For high screentime users
```{r}
df_SHMiss <- df_subset_high %>% complete(nesting(User_ID,ScreenTime), Visit)
df_SHMiss <- subset(df_SHMiss, select = c(Visit, User_ID, BL_MSEC, MSEC))
df_SHImp <- mice(df_SHMiss, m = 3, method = "rf")


complete_df_SHimp <- complete(df_SHimp, 1) #complete imputed dataest for low screentime users 
complete_df_SHimp$difference = complete_df_SHimp$MSEC - complete_df_SHimp$BL_MSEC #difference in MSEC Scores
complete_df_SHimp <- complete_df_SHimp %>% 
  group_by(User_ID) %>%
  mutate(moving_average_actual = rollapplyr(difference, 2, mean, fill=NA)) #Moving average

```


